---
title: "R Exercise 2"
author: "Azaleah Gicole"
date: "2026-09-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Part A - Arithmetic Operations

##1
```{r}
a <- 17
b <- 5

a + b
a - b
a * b
a / b
```
##2
```{r}
a <- 17
b <- 5

a %% b
a %/% b
```
##3
```{r}
2^10
```
##Q1
```{r}
17 / 5
17 %% 5
17 %/% 5
```
These three operations are related because they are a part of a division operation. The first one prints the exact quotient with a decimal, the second one prints its remainder, and the third one prints only the integer.
##Q2
```{r}
10 %% 3
10 %/% 3
```
I predicted the answer correctly, guessing 10 %% 3 prints the remainder 1 while 10 %/% 3 prints the integer 3.

## Part B - Basic Data Type
##1
```{r}
x <- 42
typeof(x)
y <- 42L
typeof(y)
z <- "42"
typeof(z)
w <- TRUE
typeof(w)
v <- 3 + 2i
typeof(v)
```
##2
```{r}
x <- 42
class(x)
y <- 42L
class(y)
z <- "42"
class(z)
w <- TRUE
class(w)
v <- 3 + 2i
class(v)
```
##3
```{r}
x <- 42
mode(x)
y <- 42L
mode(y)
z <- "42"
mode(z)
w <- TRUE
mode(w)
v <- 3 + 2i
mode(v)
```
##Q3 - x <- 42 and y <- 42L give different answers because y has the letter L in its value, turning the numeric value 42 into an integer instead.

##Q4 - Yes, typeof() gives a different output on the number values. It uses double and integer instead of numeric to check the first two classes.

## Part C - Type Conversion
##1
```{r}
num_string <- "123"
num_v <- as.numeric("num_string")
typeof(num_v)
class(num_v)
mode(num_v)
```
##2
```{r}
as.logical(0)
```
Converting 0 to 1 would result to FALSE.
##3
```{r}
bad_string <- "hello123"
good_string <- as.numeric(bad_string)
```

